The CrypTool Script: \\ Cryptography, Mathematics and More \\ \vspace{30pt} {\large Background reading\\%*[-8pt] for the free e-learning program CrypTool\\ (10th edition -- distributed with CrypTool version 1.4.30)} \vspace{60pt}

### 2009-11-21 script/en/chapters/numbertheory.tex
fixed overfull hbox
Script[de+en]: Many small corrections -- should be close to final now.

The cipher text $C$ is determined as $C = E (P; M)$, where $E$ (encryption) is the encryption rule.
The recipient uses his private key $S$ to decrypt the message with the decryption rule $D: M = D (S; C)$.

@@

\item[{\bf 6.}] For decryption, the cipher text represented as a binary number is divided
                into parts such that each part of the number is less than $n$.

\item[{\bf 7.}] Decryption of the cipher text (or the parts of it) $C \in \{1, \cdots, n-1\}$:
                $$M = D ( (n, d); C ) := C^d {\rm ~(mod~} n).$$
\end{itemize}

@@

Enciphering a text entails applying a function (mathematical operation) to a character string (number) to generate a 
different number. Deciphering entails reversing this function, in other words using the distorted image that the function 
has created from the plaintext in order to restore the original image. For example, the sender could take the plaintext 
$M$ of a confidential message and add a secret number, the key $S$, to obtain the cipher text $C$:
$$C = M + S.$$
The recipient can reconstruct the plaintext by reversing this operation, in other words by subtracting $S$:
$$M = C - S.$$
Adding $S$ reliably makes the plaintext impossible to read. However, this encryption is rather weak, because all an 
interceptor needs to do to calculate the key is obtain a plaintext and the associated cipher text
$$S = C - M,$$
and can then read any subsequent messages encrypted using $S$. \\
The essential reason for this is that subtraction is just as simple an operation as addition.

% --------------------------------------------------------------------------
\hypertarget{OneWayFunktion2}{}%
\subsection{One way functions}\index{One way function}
\label{OneWayFunktion2}%
If the key is to be impossible to determine even with knowledge of both the 
plaintext and the cipher text, we need a function that is, on the one hand, 
relatively easy to calculate -- we don't want to have problems encrypting 
messages. On the other hand, the inverse function should exist (otherwise 
information would be lost during encryption), but should be de facto 
incalculable.

@@

Before applying the RSA procedure to a text, we will first demonstrate it
directly using a single number as message\footnote{Using CrypTool\index{CrypTool} you can solve this with the menu 
{\bf Indiv. Procedures \textbackslash{} RSA Cryptosystem \textbackslash{} RSA Demonstration}.}.
\begin{itemize}
\item[{\bf 1.}] Let the selected prime numbers be $p=5$ and $q=11$. \\
Thus, $n=55$ and $J(n)=(p-1)*(q-1)=40$.
\item[{\bf 2.}] $e = 7$ (should\footnote{%
                See footnote~\ref{foot:Selection-of-e} on page
                \pageref{foot:Selection-of-e}.}
                lie between $11$ and $39$ and must be relatively prime
                to $40$).
\item[{\bf 3.}] $d = 23$ (since $23*7 \equiv 161 \equiv 1{\rm ~(mod~} 40)$),
    \begin{itemize}
    \item[] $\rightarrow$ Public key of the recipient:  $(55, 7),$
    \item[] $\rightarrow$ Private key of the recipient: $(55, 23).$
    \end{itemize}
\item[{\bf 4.}] Let the message be the number $M = 2$ (so no division into blocks is required).
\item[{\bf 5.}] Encryption: $C \equiv 2^7 \equiv 18 {\rm ~(mod~}55).$
\item[{\bf 6.}] The cipher text is simply the number $C = 18$ (we therefore do not need to divide it into blocks).
\item[{\bf 7.}] Decryption: 
        $M \equiv 18^{23} \equiv 18^{(1+2+4+16)} \equiv 18*49*36*26 \equiv 2 {\rm ~(mod~}55).$
\end{itemize}

@@

\noindent {\bf 5. Decryption}: \\
Cipher text: {\tt 1404 2932 3536 0001 3284 2280 2235 }

@@

\noindent {\bf 5. Decryption}:\\
Cipher text:

@@

\noindent {\bf 5. Decryption:} \\
Cipher text:\\
{\tt 158721 137346 37358 240130 112898}

@@

\noindent {\bf 5'. Decryption:} \\
Cipher text:\\
 {\tt 198967 051405 254571 115318 014251}

@@

Two samples of RSA cipher text are presented in Tables~\ref{stinson1}\footnote{The numbers of this table can be worked with via Copy and Paste.} and \ref{stinson2}\footnote{The numbers of this table are in the online-help ``Example illustrating the RSA demonstration'' of CrypTool\index{CrypTool}.}. 
Your task is to decrypt them. The public parameters of the system are

@@

This can be accomplished as follows. First, factor $n$ (which is easy
because it is so small). Then compute the exponent $d$ from $J(n)$, and,
finally, decrypt the cipher text. Use the square-and-multiply
\index{Square and multiply} algorithm to exponentiate modulo $n$.

@@

\begin{table}[ht]
\begin{center}
{\tt 
\begin{tabular}{llllllll}
12423 & 11524  & 7243  & 7459 & 14303  & 6127 & 10964 & 16399 \\
 9792 & 13629 & 14407 & 18817 & 18830 & 13556  & 3159 & 16647 \\
 5300 & 13951    & 81  & 8986  & 8007 & 13167 & 10022 & 17213 \\
 2264   & 961 & 17459  & 4101  & 2999 & 14569 & 17183 & 15827 \\
12693  & 9553 & 18194  & 3830  & 2664 & 13998 & 12501 & 18873 \\
12161 & 13071 & 16900  & 7233  & 8270 & 17086  & 9792 & 14266 \\
13236  & 5300 & 13951  & 8850 & 12129  & 6091 & 18110  & 3332 \\
15061 & 12347  & 7817  & 7946 & 11675 & 13924 & 13892 & 18031 \\
 2620  & 6276  & 8500   & 201  & 8850 & 11178 & 16477 & 10161 \\
 3533 & 13842  & 7537 & 12259 & 18110    & 44  & 2364 & 15570 \\
 3460  & 9886  & 8687  & 4481 & 11231  & 7547 & 11383 & 17910 \\
12867 & 13203  & 5102  & 4742  & 5053 & 15407  & 2976  & 9330 \\
12192    & 56  & 2471 & 15334   & 841 & 13995 & 17592 & 13297 \\
 2430  & 9741 & 11675   & 424  & 6686   & 738 & 13874  & 8168 \\
 7913  & 6246 & 14301  & 1144  & 9056 & 15967  & 7328 & 13203 \\
  796   & 195  & 9872 & 16979 & 15404 & 14130  & 9105  & 2001 \\
 9792 & 14251  & 1498 & 11296  & 1105  & 4502 & 16979  & 1105 \\
   56  & 4118 & 11302  & 5988  & 3363 & 15827  & 6928  & 4191 \\
 4277 & 10617   & 874 & 13211 & 11821  & 3090 & 18110    & 44 \\
 2364 & 15570  & 3460  & 9886  & 9988  & 3798  & 1158  & 9872 \\
16979 & 15404  & 6127  & 9872  & 3652 & 14838  & 7437  & 2540 \\
 1367  & 2512 & 14407  & 5053  & 1521   & 297 & 10935 & 17137 \\
 2186  & 9433 & 13293  & 7555 & 13618 & 13000  & 6490  & 5310 \\
18676  & 4782 & 11374   & 446  & 4165 & 11634  & 3846 & 14611 \\
 2364  & 6789 & 11634  & 4493  & 4063  & 4576 & 17955  & 7965 \\
11748 & 14616 & 11453 & 17666   & 925    & 56  & 4118 & 18031 \\
 9522 & 14838  & 7437  & 3880 & 11476  & 8305  & 5102  & 2999 \\
18628 & 14326  & 9175  & 9061   & 650 & 18110  & 8720 & 15404 \\
 2951   & 722 & 15334   & 841 & 15610  & 2443 & 11056  & 2186 
\end{tabular} } % tt
\end{center}
\caption{RSA cipher text A}
\label{stinson1}
\end{table}

\begin{table}[ht]
\begin{center}
{\tt 
\begin{tabular}{llllllll}
 6340  & 8309 & 14010  & 8936 & 27358 & 25023 & 16481 & 25809 \\
23614  & 7135 & 24996 & 30590 & 27570 & 26486 & 30388  & 9395 \\
27584 & 14999  & 4517 & 12146 & 29421 & 26439  & 1606 & 17881 \\
25774  & 7647 & 23901  & 7372 & 25774 & 18436 & 12056 & 13547 \\
 7908  & 8635  & 2149  & 1908 & 22076  & 7372  & 8686  & 1304 \\
 4082 & 11803  & 5314   & 107  & 7359 & 22470  & 7372 & 22827 \\
15698 & 30317  & 4685 & 14696 & 30388  & 8671 & 29956 & 15705 \\
 1417 & 26905 & 25809 & 28347 & 26277  & 7897 & 20240 & 21519 \\
12437  & 1108 & 27106 & 18743 & 24144 & 10685 & 25234 & 30155 \\
23005  & 8267  & 9917  & 7994  & 9694  & 2149 & 10042 & 27705 \\
15930 & 29748  & 8635 & 23645 & 11738 & 24591 & 20240 & 27212 \\
27486  & 9741  & 2149 & 29329  & 2149  & 5501 & 14015 & 30155 \\
18154 & 22319 & 27705 & 20321 & 23254 & 13624  & 3249  & 5443 \\
 2149 & 16975 & 16087 & 14600 & 27705 & 19386  & 7325 & 26277 \\
19554 & 23614  & 7553  & 4734  & 8091 & 23973 & 14015   & 107 \\
 3183 & 17347 & 25234  & 4595 & 21498  & 6360 & 19837  & 8463 \\
 6000 & 31280 & 29413  & 2066   & 369 & 23204  & 8425  & 7792 \\
25973  & 4477 & 30989                               
\end{tabular} } % tt
\end{center}
\caption{RSA cipher text B}
\label{stinson2}
\end{table}

@@

There are three tasks with completely different degrees of difficulty here. In
each case we know the cipher text 
and the public key $(e,n)$:
\begin{itemize}
\item[{\bf (a)}] Known plaintext: \index{Attack!known plaintext} find the secret key $d$ using the additionally known original message.
\item[{\bf (b)}] Cipher text only: \index{Attack!cipher text only} find $d$ and the plaintext.
\item[{\bf (c)}] Calculate the RSA modulus, in other words factorization (with no knowledge of the message). \index{Factorization!factorization problem}
\end{itemize}

@@

OUTPUT:
CC_Start -- Testcases for count_PrimitiveRoots_of_a_PrimesRange(start, end)
-----------Testcase: (1, 1500)
2 1
3 1
5 2
7 2
11 4
13 4
17 8
19 6
23 10
29 12
31 8
37 12
...
1483 432
1487 742
1489 480
1493 744
1499 636
Found all 62044 primitive roots of 239 primes.
Time: CPU 7.55 s, Wall: 7.85 s
\end{Verbatim}
\caption{Function ``count\_PrimitiveRoots\_of\_a\_PrimesRange'' to calculate all primitive roots for a given range of primes}
\end{sagecode}

>>>>

The ciphertext $C$ is determined as $C = E (P; M)$, where $E$ (encryption) is the encryption rule.
The recipient uses his private key $S$ to decrypt the message with the decryption rule $D: M = D (S; C)$.

@@

\item[{\bf 6.}] For decryption, the ciphertext represented as a binary number is divided
                into parts such that each part of the number is less than $n$.

\item[{\bf 7.}] Decryption of the ciphertext (or the parts of it) $C \in \{1, \cdots, n-1\}$:
                $$M = D ( (n, d); C ) := C^d {\rm ~(mod~} n).$$
\end{itemize}

@@

Enciphering a text entails applying a function (mathematical operation) to a character string (number) to generate a 
different number. Deciphering entails reversing this function, in other words using the distorted image that the function 
has created from the plaintext in order to restore the original image. For example, the sender could take the plaintext 
$M$ of a confidential message and add a secret number, the key $S$, to obtain the ciphertext $C$:
$$C = M + S.$$
The recipient can reconstruct the plaintext by reversing this operation, in other words by subtracting $S$:
$$M = C - S.$$
Adding $S$ reliably makes the plaintext impossible to read. However, this encryption is rather weak, because all an 
interceptor needs to do to calculate the key is obtain a plaintext and the associated ciphertext
$$S = C - M,$$
and can then read any subsequent messages encrypted using $S$. \\
The essential reason for this is that subtraction is just as simple an operation as addition.

% --------------------------------------------------------------------------
\hypertarget{OneWayFunktion2}{}%
\subsection{One way functions}\index{One way function}
\label{OneWayFunktion2}%
If the key is to be impossible to determine even with knowledge of both the 
plaintext and the ciphertext, we need a function that is, on the one hand, 
relatively easy to calculate -- we don't want to have problems encrypting 
messages. On the other hand, the inverse function should exist (otherwise 
information would be lost during encryption), but should be de facto 
incalculable.

@@

Before applying the RSA procedure to a text, we will first demonstrate it
directly using a single number as message\footnote{Using CrypTool\index{CrypTool} you can solve this with the menu 
{\bf Indiv. Procedures \textbackslash{} RSA Cryptosystem \textbackslash{} RSA Demonstration}.}.
\begin{itemize}
\item[{\bf 1.}] Let the selected prime numbers be $p=5$ and $q=11$. \\
Thus, $n=55$ and $J(n)=(p-1)*(q-1)=40$.
\item[{\bf 2.}] $e = 7$ (should\footnote{%
                See footnote~\ref{foot:Selection-of-e} on page
                \pageref{foot:Selection-of-e}.}
                lie between $11$ and $39$ and must be relatively prime
                to $40$).
\item[{\bf 3.}] $d = 23$ (since $23*7 \equiv 161 \equiv 1{\rm ~(mod~} 40)$),
    \begin{itemize}
    \item[] $\rightarrow$ Public key of the recipient:  $(55, 7),$
    \item[] $\rightarrow$ Private key of the recipient: $(55, 23).$
    \end{itemize}
\item[{\bf 4.}] Let the message be the number $M = 2$ (so no division into blocks is required).
\item[{\bf 5.}] Encryption: $C \equiv 2^7 \equiv 18 {\rm ~(mod~}55).$
\item[{\bf 6.}] The ciphertext is simply the number $C = 18$ (we therefore do not need to divide it into blocks).
\item[{\bf 7.}] Decryption: 
        $M \equiv 18^{23} \equiv 18^{(1+2+4+16)} \equiv 18*49*36*26 \equiv 2 {\rm ~(mod~}55).$
\end{itemize}

@@

\noindent {\bf 5. Decryption}: \\
Ciphertext: {\tt 1404 2932 3536 0001 3284 2280 2235 }

@@

\noindent {\bf 5. Decryption}:\\
Ciphertext:

@@

\noindent {\bf 5. Decryption:} \\
Ciphertext:\\
{\tt 158721 137346 37358 240130 112898}

@@

\noindent {\bf 5'. Decryption:} \\
Ciphertext:\\
 {\tt 198967 051405 254571 115318 014251}

@@

Two samples of RSA ciphertext are presented in Tables~\ref{stinson1}\footnote{The numbers of this table can be worked with via Copy and Paste.} and \ref{stinson2}\footnote{The numbers of this table are in the online-help ``Example illustrating the RSA demonstration'' of CrypTool\index{CrypTool}.}. 
Your task is to decrypt them. The public parameters of the system are

@@

This can be accomplished as follows. First, factor $n$ (which is easy
because it is so small). Then compute the exponent $d$ from $J(n)$, and,
finally, decrypt the ciphertext. Use the square-and-multiply
\index{Square and multiply} algorithm to exponentiate modulo $n$.

@@

\begin{table}[ht]
\begin{center}
{\tt 
\begin{tabular}{llllllll}
12423 & 11524  & 7243  & 7459 & 14303  & 6127 & 10964 & 16399 \\
 9792 & 13629 & 14407 & 18817 & 18830 & 13556  & 3159 & 16647 \\
 5300 & 13951    & 81  & 8986  & 8007 & 13167 & 10022 & 17213 \\
 2264   & 961 & 17459  & 4101  & 2999 & 14569 & 17183 & 15827 \\
12693  & 9553 & 18194  & 3830  & 2664 & 13998 & 12501 & 18873 \\
12161 & 13071 & 16900  & 7233  & 8270 & 17086  & 9792 & 14266 \\
13236  & 5300 & 13951  & 8850 & 12129  & 6091 & 18110  & 3332 \\
15061 & 12347  & 7817  & 7946 & 11675 & 13924 & 13892 & 18031 \\
 2620  & 6276  & 8500   & 201  & 8850 & 11178 & 16477 & 10161 \\
 3533 & 13842  & 7537 & 12259 & 18110    & 44  & 2364 & 15570 \\
 3460  & 9886  & 8687  & 4481 & 11231  & 7547 & 11383 & 17910 \\
12867 & 13203  & 5102  & 4742  & 5053 & 15407  & 2976  & 9330 \\
12192    & 56  & 2471 & 15334   & 841 & 13995 & 17592 & 13297 \\
 2430  & 9741 & 11675   & 424  & 6686   & 738 & 13874  & 8168 \\
 7913  & 6246 & 14301  & 1144  & 9056 & 15967  & 7328 & 13203 \\
  796   & 195  & 9872 & 16979 & 15404 & 14130  & 9105  & 2001 \\
 9792 & 14251  & 1498 & 11296  & 1105  & 4502 & 16979  & 1105 \\
   56  & 4118 & 11302  & 5988  & 3363 & 15827  & 6928  & 4191 \\
 4277 & 10617   & 874 & 13211 & 11821  & 3090 & 18110    & 44 \\
 2364 & 15570  & 3460  & 9886  & 9988  & 3798  & 1158  & 9872 \\
16979 & 15404  & 6127  & 9872  & 3652 & 14838  & 7437  & 2540 \\
 1367  & 2512 & 14407  & 5053  & 1521   & 297 & 10935 & 17137 \\
 2186  & 9433 & 13293  & 7555 & 13618 & 13000  & 6490  & 5310 \\
18676  & 4782 & 11374   & 446  & 4165 & 11634  & 3846 & 14611 \\
 2364  & 6789 & 11634  & 4493  & 4063  & 4576 & 17955  & 7965 \\
11748 & 14616 & 11453 & 17666   & 925    & 56  & 4118 & 18031 \\
 9522 & 14838  & 7437  & 3880 & 11476  & 8305  & 5102  & 2999 \\
18628 & 14326  & 9175  & 9061   & 650 & 18110  & 8720 & 15404 \\
 2951   & 722 & 15334   & 841 & 15610  & 2443 & 11056  & 2186 
\end{tabular} } % tt
\end{center}
\caption{RSA ciphertext A}
\label{stinson1}
\end{table}

\begin{table}[ht]
\begin{center}
{\tt 
\begin{tabular}{llllllll}
 6340  & 8309 & 14010  & 8936 & 27358 & 25023 & 16481 & 25809 \\
23614  & 7135 & 24996 & 30590 & 27570 & 26486 & 30388  & 9395 \\
27584 & 14999  & 4517 & 12146 & 29421 & 26439  & 1606 & 17881 \\
25774  & 7647 & 23901  & 7372 & 25774 & 18436 & 12056 & 13547 \\
 7908  & 8635  & 2149  & 1908 & 22076  & 7372  & 8686  & 1304 \\
 4082 & 11803  & 5314   & 107  & 7359 & 22470  & 7372 & 22827 \\
15698 & 30317  & 4685 & 14696 & 30388  & 8671 & 29956 & 15705 \\
 1417 & 26905 & 25809 & 28347 & 26277  & 7897 & 20240 & 21519 \\
12437  & 1108 & 27106 & 18743 & 24144 & 10685 & 25234 & 30155 \\
23005  & 8267  & 9917  & 7994  & 9694  & 2149 & 10042 & 27705 \\
15930 & 29748  & 8635 & 23645 & 11738 & 24591 & 20240 & 27212 \\
27486  & 9741  & 2149 & 29329  & 2149  & 5501 & 14015 & 30155 \\
18154 & 22319 & 27705 & 20321 & 23254 & 13624  & 3249  & 5443 \\
 2149 & 16975 & 16087 & 14600 & 27705 & 19386  & 7325 & 26277 \\
19554 & 23614  & 7553  & 4734  & 8091 & 23973 & 14015   & 107 \\
 3183 & 17347 & 25234  & 4595 & 21498  & 6360 & 19837  & 8463 \\
 6000 & 31280 & 29413  & 2066   & 369 & 23204  & 8425  & 7792 \\
25973  & 4477 & 30989                               
\end{tabular} } % tt
\end{center}
\caption{RSA ciphertext B}
\label{stinson2}
\end{table}

@@

There are three tasks with completely different degrees of difficulty here.
In each case we know the ciphertext 
and the public key $(e,n)$:
\begin{itemize}
\item[{\bf (a)}] Known plaintext: \index{Attack!known plaintext} find the secret key $d$ using the additionally known original message.
\item[{\bf (b)}] Ciphertext-only: \index{Attack!ciphertext-only} find $d$ and the plaintext.
\item[{\bf (c)}] Calculate the RSA modulus, in other words factorization (with no knowledge of the message). \index{Factorization!factorization problem}
\end{itemize}

@@

OUTPUT:
CC_Start -- Testcases for count_PrimitiveRoots_of_a_PrimesRange(start, end)
-----------Testcase: (1, 1500)
2 1
3 1
5 2
7 2
11 4
13 4
17 8
19 6
23 10
29 12
31 8
37 12
...
1483 432
1487 742
1489 480
1493 744
1499 636
Found all 62044 primitive roots of 239 primes.
Time: CPU 7.55 s, Wall: 7.85 s
\end{Verbatim}
\caption{Function ``count\_PrimitiveRoots\_of\_a\_PrimesRange'' to calculate the number of primitive roots for a given range of primes}
\end{sagecode}

### 2009-11-21 script/en/chapters/paper_and_pencil.tex
Script[de+en]: Many small corrections -- should be close to final now.

In the following section some classic ciphers are implemented using the open source computer
algebra system Sage. The code was tested with Sage version 4.2.
All ciphers are explained in chapter~\ref{Kapitel_PaperandPencil}
(``\nameref{Kapitel_PaperandPencil}'').

@@

In the above samples we applied the SubstitutionCryptosystem and build
Caesar as a special kind of substitution. In contrast here
Caesar can be build as a special kind of the shift cipher.

@@

In the following Sage example the symbols are from the binary number
system. A monoalphabetic substitution cipher with a binary alphabet has very little
security: Because the plaintext/ciphertext alphabet has only the two elements
0 and 1, there are only two keys possible: (0 1) and (1 0). Remark: In each key of a
substitution cipher all symbols of the alphabet have to appear once.

@@

@@

\bibitem[Crowley2000]{pp:Crowley2000} \index{Crowley 2000}
   Paul Crowley, \\
   {\em Mirdek: A card cipher inspired by ``Solitaire''}, \\
   2000.\\
   \href{http://www.ciphergoth.org/crypto/mirdek/}
   {\texttt{http://www.ciphergoth.org/crypto/mirdek/}}

\bibitem[DA1999]{pp:DA1999} \index{DA 1999}
   Data encryption page of the ThinkQuest Team 27158 for ThinkQuest 1999 \\
   (no update since 1999, no search possibility), \\
   1999.\\
   \href{http://library.thinkquest.org/27158/}
   {\texttt{http://library.thinkquest.org/27158/}}

\bibitem[Goebel2003]{pp:Goebel2003} \index{Goebel 2003}
   Greg Goebel, \\
   {\em Codes, Ciphers and Codebreaking}, \\
   2003.\\
   \href{http://www.vectorsite.net/ttcode.htm}
   {\texttt{http://www.vectorsite.net/ttcode.htm}}

@@

\bibitem[Nichols1996]{pp:Nichols1996} \index{Nichols 1996} 
    Randall K. Nichols, \\
    {\em Classical Cryptography Course, Volume 1 and 2}, \\
    Aegean Park Press 1996;
    or in 12 lessons online at \\
    \href{http://www.fortunecity.com/skyscraper/coding/379/lesson1.htm}
    {\texttt{http://www.fortunecity.com/skyscraper/coding/379/lesson1.htm}}

\bibitem[Savard1999]{pp:Savard1999} \index{Savard 1999}
	John J. G. Savard, \\
	{\em A Cryptographic Compendium}, \\
	1999.\\
	\href{http://www.hypermaths.org/quadibloc/crypto/jscrypt.htm}
	{\texttt{http://www.hypermaths.org/quadibloc/crypto/jscrypt.htm}}

@@

\bibitem[ThinkQuest1999]{pp:ThinkQuest1999} \index{ThinkQuest 1999}
	ThinkQuest Team 27158, \\
	{\em Data Encryption}, \\
	1999.\\
	\href{http://library.thinkquest.org/27158/}
	{\texttt{http://library.thinkquest.org/27158/} }

>>>>

In the following section some classic ciphers are implemented using the open source computer
algebra system (CAS) \verb!Sage!. The code was tested with Sage version 4.2.
All ciphers are explained in chapter~\ref{Kapitel_PaperandPencil}
(``\nameref{Kapitel_PaperandPencil}'').

@@

In the above samples we applied the \verb!SubstitutionCryptosystem!
and build Caesar as a special kind of substitution. In contrast here
Caesar can be build as a special kind of the shift cipher.

@@

In the following Sage example the symbols are from the binary number
system. A monoalphabetic substitution cipher with a binary alphabet has very little
security: Because the plaintext/ciphertext alphabet has only the two elements
0 and 1, there are only two keys possible: (0 1) and (1 0). \\
Remark: The key of a general substitution cipher contains all symbols
of the alphabet exactly once.

@@

Remark: Currently \verb!S! has no attribute \verb!key!, and I found no way
to transform the binary sequence \verb!DC! back to ASCII.

@@

\bibitem[Crowley2000]{pp:Crowley2000} \index{Crowley 2000}
   Paul Crowley, \\
   {\em Mirdek: A card cipher inspired by ``Solitaire''}, \\
   2000.\\
   \url{http://www.ciphergoth.org/crypto/mirdek/}

\bibitem[DA1999]{pp:DA1999} \index{DA 1999}
   Data encryption page of the ThinkQuest Team 27158 for ThinkQuest 1999 \\
   (no update since 1999, no search possibility), \\
   1999.\\
   \url{http://library.thinkquest.org/27158/}

\bibitem[Goebel2003]{pp:Goebel2003} \index{Goebel 2003}
   Greg Goebel, \\
   {\em Codes, Ciphers and Codebreaking}, \\
   2003.\\
   \url{http://www.vectorsite.net/ttcode.htm}

@@

\bibitem[Nichols1996]{pp:Nichols1996} \index{Nichols 1996} 
    Randall K. Nichols, \\
    {\em Classical Cryptography Course, Volume 1 and 2}, \\
    Aegean Park Press 1996;
    or in 12 lessons online at \\
    \url{http://www.fortunecity.com/skyscraper/coding/379/lesson1.htm}

\bibitem[Savard1999]{pp:Savard1999} \index{Savard 1999}
	John J. G. Savard, \\
	{\em A Cryptographic Compendium}, \\
	1999.\\
	\url{http://www.quadibloc.com/crypto/jscrypt.htm}

@@

\bibitem[ThinkQuest1999]{pp:ThinkQuest1999} \index{ThinkQuest 1999}
	ThinkQuest Team 27158, \\
	{\em Data Encryption}, \\
	1999.\\
	\url{http://library.thinkquest.org/27158/}
